Classless Java: Tuning Java Interfaces

### 2015-12-08 paper/paper-without-appendix.tex
clean up and updated pdfs

\begin{abstract}

Java 8 introduced \emph{default methods}, allowing interfaces to
have method implementations. When combined with (multiple) interface
inheritance, default methods provide a basic form of multiple
inheritance. However, using this combination to simulate multiple
inheritance quickly becomes cumbersome, and appears to be quite
restricted.

This paper shows that, with a simple language feature, default methods
and interface inheritance are in fact very expressive. Our proposed
language feature, called \emph{object interfaces}, enables powerful
object-oriented idioms, using multiple inheritance, to be expressed
conveniently in Java. Object interfaces refine conventional Java
interfaces in three different ways. Firstly, object interfaces have
their own object instantiation mechanism, providing an alternative to
class constructors. Secondly, object interfaces support \emph{abstract
  state operations}, providing a way to use multiple inheritance with
state in Java. Finally, object interfaces allow type refinements that
are often tricky to model in conventional class-based
approaches. Interestingly, object interfaces do not require changes to
the runtime, and they also do not introduce any new syntax: all three
features are achieved by reinterpreting existing Java syntax, and are
translated into regular Java code without loss of type-safety. Since
no new syntax is introduced, it would be incorrect to call object
interfaces a language extension or syntactic sugar. So we use the term
\emph{language tuning} to characterize this kind of language
feature. An implementation of object interfaces using Java annotations
and a formalization of the static and dynamic semantics are
presented. Moreover the usefulness of object interfaces is illustrated
through various examples.

\end{abstract}

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

@@

\appendix
\input{sections/Appendix.tex}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "."
%%% End:

>>>>

\input{sections/Abstract.tex}

@@

@@

### 2015-12-08 paper/sections/Abstract.tex
clean up and updated pdfs

%\documentclass[preprint]{sigplanconf}
%\documentclass[preprint]{llncs}
\documentclass[a4paper,UKenglish]{lipics}

\input{def}

\usepackage{listings}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
\usepackage[table,xcdraw]{xcolor}
\usepackage{color}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
%% \usepackage{mathpartir}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}

\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{mathpartir}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{thm2}{Theorem}
\newtheorem{lem2}{Lemma}

% "define" code highlights for Java and Scala
\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,void,catch,class,def,static,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,var,while,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=3pt,
belowskip=-1pt
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

%%\usepackage{natbib}
%%\bibpunct();A{},
%%\let\cite=\citep

%include lhs2TeX.fmt
%include lhs2TeX.sty
%include forall.fmt

%\pagestyle{plain}

%{\theorembodyfont{\sffamily} \newtheorem{theorem}{Theorem}}
%{\theorembodyfont{\sffamily} \newtheorem{lemma}{Lemma}}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}

%%%%========== authornotes marco begin ================
%%use the following 5 lines for showing notes
% \newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
% \newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
% \newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
% \newcommand\marco[1]{\authornote{marco}{blue}{#1}}
% \newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
%%use the following 4 lines for hiding notes
\newcommand\bruno[1]{}
\newcommand\yanlin[1]{}
\newcommand\marco[1]{}
\newcommand\haoyuan[1]{}
%%%%========== authornotes marco end ====================

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\newcommand\mixin{\mixinAnn\xspace}
\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}
\usepackage{xspace}

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Classless Java: Tuning Java Interfaces}
\titlerunning{Classless Java}

\author[1]{}
% \author[2]{Joan R. Access}
% \affil[1]{Dummy University Computing Laboratory\\
%   Address, Country\\
%   \texttt{open@dummyuni.org}}
% \affil[2]{Department of Informatics, Dummy College\\
%   Address, Country\\
%   \texttt{access@dummycollege.org}}
% \authorrunning{J.\,Q. Open and J.\,R. Access}
% mandatory. First: Use abbreviated first/middle names. Second (only in severe
% cases): Use first author plus 'et. al.'

\begin{comment}
\Copyright{John Q. Open and Joan R. Access}
% mandatory, please use full first names. LIPIcs license is "CC-BY";
% http://creativecommons.org/licenses/by/3.0/

\subjclass{Dummy classification -- please refer to
  \url{http://www.acm.org/about/class/ccs98-html}}
% mandatory: Please choose ACM 1998 classifications from
% http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of
% Computation".
\keywords{Dummy keyword -- please provide 1--5 keywords}% mandatory: Please provide 1-5 keywords
% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\serieslogo{}%please provide filename (without suffix)
\volumeinfo%(easychair interface)
  {Billy Editor and Bill Editors}% editors
  {2}% number of editors: 1, 2, ....
  {Conference title on which this volume is based on}% event
  {1}% volume
  {1}% issue
  {1}% starting page number
\EventShortName{}
\DOI{10.4230/LIPIcs.xxx.yyy.p}% to be completed by the volume editor
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle

\begin{abstract}

Java 8 introduced \emph{default methods}, allowing interfaces to
have method implementations. When combined with (multiple) interface
inheritance, default methods provide a basic form of multiple
inheritance. However, using this combination to simulate multiple
inheritance quickly becomes cumbersome, and appears to be quite
restricted.

This paper shows that, with a simple language feature, default methods
and interface inheritance are in fact very expressive. Our proposed
language feature, called \emph{object interfaces}, enables powerful
object-oriented idioms, using multiple inheritance, to be expressed
conveniently in Java. Object interfaces refine conventional Java
interfaces in three different ways. Firstly, object interfaces have
their own object instantiation mechanism, providing an alternative to
class constructors. Secondly, object interfaces support \emph{abstract
  state operations}, providing a way to use multiple inheritance with
state in Java. Finally, object interfaces allow type refinements that
are often tricky to model in conventional class-based
approaches. Interestingly, object interfaces do not require changes to
the runtime, and they also do not introduce any new syntax: all three
features are achieved by reinterpreting existing Java syntax, and are
translated into regular Java code without loss of type-safety. Since
no new syntax is introduced, it would be incorrect to call object
interfaces a language extension or syntactic sugar. So we use the term
\emph{language tuning} to characterize this kind of language
feature. An implementation of object interfaces using Java annotations
and a formalization of the static and dynamic semantics are
presented. Moreover the usefulness of object interfaces is illustrated
through various examples.

\end{abstract}

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

%===============================================================================

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Formalization.tex}

\input{sections/Translation.tex}

\input{sections/Implementation.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/FutureWork.tex}

%===============================================================================
\input{sections/Conclusion.tex}

%===============================================================================

%\bibliographystyle{splncs}
\bibliographystyle{plain}
\bibliography{paper}

\appendix
\input{sections/Appendix.tex}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "."
%%% End:

>>>>

\begin{abstract}
  Java 8 introduced \emph{default methods}, allowing interfaces to have method
  implementations. When combined with (multiple) interface inheritance, default
  methods provide a basic form of multiple inheritance. However, using this
  combination to simulate multiple inheritance quickly becomes cumbersome, and
  appears to be quite restricted.

  This paper shows that, with a simple language feature, default methods and
  interface inheritance are in fact very expressive. Our proposed language
  feature, called \emph{object interfaces}, enables powerful object-oriented
  idioms, using multiple inheritance, to be expressed conveniently in
  Java. Object interfaces refine conventional Java interfaces in three different
  ways. Firstly, object interfaces have their own object instantiation
  mechanism, providing an alternative to class constructors. Secondly, object
  interfaces support \emph{abstract state operations}, providing a way to use
  multiple inheritance with state in Java. Finally, object interfaces allow type
  refinements that are often tricky to model in conventional class-based
  approaches. Interestingly, object interfaces do not require changes to the
  runtime, and they also do not introduce any new syntax: all three features are
  achieved by reinterpreting existing Java syntax, and are translated into
  regular Java code without loss of type-safety. Since no new syntax is
  introduced, it would be incorrect to call object interfaces a language
  extension or syntactic sugar. So we use the term \emph{language tuning} to
  characterize this kind of language feature. An implementation of object
  interfaces using Java annotations and a formalization of the static and
  dynamic semantics are presented. Moreover the usefulness of object interfaces
  is illustrated through various examples.
\end{abstract}
